---
title: "Econometric Studies in Brazil"
author: "claudio freitas"
date: "Saturday, September 12, 2015"
output: word_document
---

##EXECUTIVE SUMMARY
In this project I am using the QUANDL database to analyse some current Brazilian economic data in order to try to predict short term path of inflation and how long should we wait until we can see growth back to the Brazilian economy.


#Data Processing

Dowloading the data from QUAnDL.

```{r, echo=TRUE}
library(Quandl)
library(tseries)
library(timeSeries)
library(forecast)
library("zoo")
Quandl.auth("xsPzjKeHjtLjjjEpLgYi")
PrimeBR<- Quandl("BCB/20019")
InfBR<- Quandl("BCB/191")
GDPBRraw<-Quandl("IBGE/ST12_BR_BRASIL_PERC")
str(GDPBRraw)
#do not consider inflation befre 1996
newinf1<- subset(InfBR, InfBR$Date > "2005-03-31")
newinf<- subset(newinf1, newinf1$Date < "2015-07-31")
#print(GDPBRraw)
#print(PrimeBR)

#print(newinf)
n<- dim(newinf)[1]/3

inf<- data.frame(matrix(0,n))
i2<- 1
for (i in 1:n)   {
          
          inf[i,1]<- (((newinf$Value[i2]/100+1)*(newinf$Value[i2+1]/100+1)*
                     (newinf$Value[i2+2]/100+1))-1)*100
          i2<- i2+3
     }

#print(inf)
inf<- apply(inf,2, rev)
combo1<- merge(InfBR, PrimeBR, by="Date")
combo<- merge(combo1,GDPBRraw, by="Date" )
combo<- cbind(combo, inf)
print(combo)
```

Test to find out if data is stationary:

```{r, echo=TRUE}
#as p value is greater than 0.05 data is non stationary
test1<-adf.test(combo$matrix.0..n., alternative= "stationary")
print(test1)
plot(combo$matrix.0..n., type="l")
test2<-adf.test(combo$Value.y, alternative= "stationary")
print(test2)

test3<-adf.test(combo$Value, alternative= "stationary")
print(test3)

```

As the quarterly inflation data seems to be non stationary, I will differentiate the data.


```{r, echo=TRUE}
nd<-ndiffs(combo$matrix.0..n.)
print(nd)
primediff<- diff(combo$matrix.0..n., differences=nd)
combo<- combo[-1, ]
combo<-cbind(combo, primediff)
print(combo)
plot(combo$primediff, type="l")
test4<-adf.test(combo$primediff, alternative= "stationary")
print(test4)
```

#Econometric models

Lets try the auto arima function to predict inflation based on past behaviour and prime rate:

```{r, echo=TRUE}
fit<- auto.arima(combo$primediff, xreg=combo$Value.y, stepwise=FALSE, approximation=FALSE)
print(fit)
```

Forecast inflation differential for the next 6 quarters:

```{r, echo=TRUE}
x<-c(17, 17, 17, 17, 17, 17)
fore6<-forecast(fit, h=6, xreg = x)
plot(fore6)
fore6<- data.frame(fore6)
last<- tail(combo$matrix.0..n.,1)
print(last)
print(fore6)
```

Calculating the forecats for the inflation quarter by quarter:

```{r, echo=F}
last1<-(fore6[1,1]+last)
last2<-(fore6[2,1]+last1)
last3<-(fore6[3,1]+last2)
last4<-(fore6[4,1]+last3)
last5<-(fore6[5,1]+last4)
last6<-(fore6[6,1]+last5)
print("Forecast quarterly inflation:")
print(last1)
print(last2)
print(last3)
print(last4)
print(last5)
print(last6)
```

Lets try another auto arima function to predict quarterly GDP based on past behaviour and prime rate:


```{r, echo=TRUE}
fit<- auto.arima(combo$Value, xreg=combo$Value.y, stepwise=FALSE, approximation=FALSE)
print(fit)
```

Forecast GDP for the next 6 quarters:

```{r, echo=TRUE}
fore62<-forecast(fit, h=6, xreg = x)
plot(fore62)
fore62<- data.frame(fore62)
print(fore62)
```

#Conclusion

Even thought the confidence intervals are pretty big, we can learn some lessons in this econometric study. First is that the outlook for inflation in Brazil in the following quarters remains still uncertain. Despite the short term good news, inflation could rebound in the endof the year and interest rates will have to remain pretty high during 2016 to prevent new pressures in the following quarters.

Regarding the GDP, the outlook remains negative. Some positive figures can appear in the following months, but recovery seems to be slow and painful. It seems that recession in Brazil will continue throughout 2016.
